---
title: "Germination of Cistus seeds depending on temperature"
author: "Gaillard Clément, Hamburger Malo, Perez Quentin"
date: "April 10th 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Define here your working directory (where data are stored)
knitr::opts_knit$set(root.dir="C:/Users/Malo Hamburger/Documents/R/TP Orgamil")

# Here are some specific packages you can use for analyses
require(lme4) # To perform mixed-effect models (see below)
require(ggplot2) # For figures
```

```{r data}
# Get the complete dataset
data <- read.table("Germination cistes_2020.csv", sep=";", 
                   dec=",", header=T, stringsAsFactors = F)[1:882,1:8]

data$Year <- as.factor(data$Year)

# For analysis of calendar data
data$Date <- as.Date(data$Date, format="%d/%m/%Y")
# ID for each seed germination assay
data$Exp <- paste(data$Group,data$Temperature,data$Replicate, sep="_")

# Numerical version of temperatures
data$Temp.num <- data$Temperature
data$Temp.num[data$Temp.num=="Controle"] <- "20"
data$Temp.num <- as.numeric(data$Temp.num)

# Have a look at the number of values per species and per temperatyre
table(data$Temperature, data$Species)

# The groups who provided the data
unique(data$Group)
# 35 groups have provided results
```

## Introduction

*Explain here the context, the general objectives, and the specific hypotheses that are tested in the study*

## Material and Methods

### Germination experiment

*Shortly present the selected species and the origin of seeds*

### Germination experiment

*Describe here the experimental design*

### Statistical analyses

We analyzed the number of seeds that have germinated at the end of experiment.

```{r data.end}
# The new table will only include the number of germinated seeds overall for each assay
data.end <- data.frame(row.names=sort(unique(data$Exp)))

# Germination estimated for 1 replicate = 15 seeds
data.end$Total.germination <- tapply(data$Germinated.seeds.per.replicate, data$Exp, sum)

# Information on the assays
data.end$Species <- tapply(data$Species, data$Exp, unique)
data.end$Year <- tapply(data$Year, data$Exp, unique)
data.end$Group <- tapply(data$Group, data$Exp, unique)
data.end$Temperature <- tapply(data$Temperature, data$Exp, unique)
data.end$Temp.num <- tapply(data$Temp.num, data$Exp, unique)
data.end$Replicate <- tapply(data$Replicate, data$Exp, unique)
```

We can note that the distribution of germinated seeds is very different for *monspeliensis* in 2019 and 2020. Almost no seeds germinated in 2019, while many did in 2020.
In 2020, the seeds were sampled in a new population.

```{r monspeliensis.years}
 h <- ggplot(data = data.end[data.end$Species=="monspeliensis",], aes(x=Total.germination, fill=Year)) +
    geom_histogram(bins=15, alpha=0.5) +
    labs(fill="")
print(h)
```

For subsequent analyses, we only keep the data on *monspeliensis* in 2020.

```{r monspeliensis.sel}
data.end <- data.end[!(data.end$Species=="monspeliensis" & data.end$Year=="2019"),]
```

## Results

```{r, echo=FALSE}
# Have a look at the data
View(data.end)
```

```{r fig.germin}
# See the distribution of germinated seeds at the end of experiment, for all species
hist(data.end$Total.germination, main="Total number of germinated seeds at the end of experiment", xlab="Number of germinated seeds")

# Comparing species germination at each temperature
par(mfrow=c(2,2))
for(treatment in unique(data.end$Temperature))
{
    subset <- (!is.na(data.end$Total.germination) & data.end$Temperature==treatment)
    boxplot(Total.germination~Species, data = data.end[subset,], main=treatment)
}

# Germination of each species at different temperatures
par(mfrow=c(1,3))
for(species in unique(data.end$Species))
{
  subset <- (!is.na(data.end$Total.germination) & data.end$Species==species)
  boxplot(Total.germination~Temp.num, data = data.end[subset,], main=species, ylim=c(0,14))
}

```

Now let us design a linear model to statistically test the influence of temperature and species on the number of germinated seeds.
 
We start with a basic linear model. We hypothesize that germination is a quadratic function of temperature (with a bell-shaped curve). To represent this effect, you must include as predictors temperature and temperature^2.

```{r simple.lm}
# Statistical test of the influence of temperature and species
germ.lm <- lm(###COMPLETE HERE###)
summary(germ.lm)
```

Let us have a look at the residuals.

```{r lm.resid}
par(mfrow=c(2,2))
plot(germ.lm)
```

*Please comment and conclude on the validity of this model.*

What about the distribution of seed number for each treatment?
Is it fine to perform a standard linear model on such data?

```{r, distrib.albidus}
sp <- "albidus"
temp <- "100"
hist(data.end$Total.germination[data.end$Species==sp & data.end$Temp.num==temp])
```

*Please suggest data transformation to manage the asymmetry of this distribution, and perform a new model*

```{r}
###COMPLETE HERE###
```

*Are we happy with the new results?*

In fact you remember that you have integer number of germinated seeds, which are unlikely to follow a Gaussian distribution?
This kind of count data should follow a Poisson distribution.
Luckily, you have a method allowing to analyze results for such data, the Poisson regression.

```{r glm.poisson}
# GLM with Poisson distribution
germ.glm <- glm(###COMPLETE HERE###, family="poisson")
summary(germ.glm)
plot(germ.glm)
```

*Is it ok now?*

In fact there is another issue: although the groups should apply the same protocol, there could be some subtle differences in the way the did the job (different way to clean the seeds and putting them in boxes...).
We need to control the fact that groups did not apply exactly the same protocol. To do that, you can incorporate the influence of group in the model.

```{r glmer.poisson}
germ.lmer <- lmer(Total.germination~(Temp.num + I(Temp.num^2))*Species + (1|Group), data=data.end)
germ.lmer0 <- lmer(Total.germination~Species + (1|Group), data=data.end)
germ.lmer1 <- lmer(Total.germination~Temp.num + I(Temp.num^2) + (1|Group), data=data.end)
germ.lmer2 <- lmer(Total.germination~Temp.num + I(Temp.num^2) + Species + (1|Group), data=data.end)
anova(germ.lmer, germ.lmer0)
anova(germ.lmer, germ.lmer1)
anova(germ.lmer, germ.lmer2)
```

Now you should have designed an appropriate model. :-)

## Discussion

*Synthesize here the main lessons of the study*

*Expose the limitation of the experiment and of the analyses, and propose some perspectives to investigate further the issue*

## References
